Modular Model-Checking for Fault-Tolerant Distributed System Implementations

### 2016-12-07 doc/om-paper/paper.tex
merged conflict.
Fix screwup in the experiments table
Cleanup and making a little space in paper

\newboolean{submission}  %set to true for the submission version
\setboolean{submission}{false}
%\setboolean{submission}{true}
\ifthenelse
{\boolean{submission}}
{ %
  \newcommand{\lee}[1]{ } %
  \newcommand{\ben}[1]{ } %
  % Put your own TODO macros here and below
} %hide todo
{ %
  \newcommand{\lee}[1]{ {\color{blue}$<$lee: #1$>$} } %
  \newcommand{\ben}[1]{ {\color{purple}$<$ben: #1$>$} } %
  \usepackage{hyperref} %
  \usepackage[inline]{showlabels} %
}

@@

%\author{Benjamin F Jones and Lee Pike}
\author{Benjamin F Jones\inst{1} \and Lee Pike\inst{2}}

@@

\lee{Ben will redo. I'd just remove the Gurevich reference}

@@

\begin{example}
Consider a simple distributed system involving 3 processors: $S$, $L_1$, and $L_2$. The purpose of the system is for $S$ to broadcast a message to $L_1$ and $L_2$ after which they perform a computation on the message and update their local state with the result. One might model this system as a transition system whose state is the product of the local processor states along with the state of the environment and any globally important phenomena such as time.

\begin{figure}
\centering
\begin{tikzpicture}[->,>=stealth',shorten >=1pt,auto,node distance=3cm,
  thick,main node/.style={circle,fill=blue!20,draw,
  font=\sffamily\Large\bfseries,minimum size=8mm}]

  \node[main node] (S) at (0,0) {$S$};
  \node[main node] (L1) at (2,1) {$L_1$};
  \node[main node] (L2) at (2,-1) {$L_2$};

  \path[every node/.style={font=\sffamily\small,
          fill=white,inner sep=1pt}]
    (S)  edge (L1)
         edge (L2);

\end{tikzpicture}
\label{fig:exa-trans}
\caption{Simple message passing system}
\end{figure}

Intuitively, there are subsets of the overall state space which are important to understanding the evolution of such a system. For example, let $A_1$ denote the subset of states in which the processor $S$ has not sent a message, let $A_2$ be the subset in which $S$ has broadcast a message, and let $A_3$ be the subset in which $L_1$ and $L_2$ have each received and processed the message. The union $A_1 \cup A_2 \cup A_3$ covers the set of reachable states of the system and defines an ASM that is useful for reasoning about its dynamics and understanding counter-example traces.
\end{example}

% ------------------------------------------------------------

% ------------------------------------------------------------
\section{Putting it Together: Compositional Modeling and Verification for Oral Messages}\label{sec:byz}

@@

\begin{figure}
  \centering
  \begin{tabular}{c|c||c|c|c|c|c|c|c|c|c|c|}
      \multicolumn{2-12}{c}{Receivers} \\
      \cline{2-12}
          &     & 1  &  2  &   3  &  4  &   5  &  6   &  7   &  8  &  9  &  10 \\
      \cline{2-12} \cline{2-12}
      \multirow{10}{*}{Relays} &
            1   & 7    & 9   & 12   & 15  & 21   & 25   & 32   & 40  & 54  & 74  \\
      \cline{2-12}
          & 2   & 17   & 14  & 21   & 30  & 42   & 53   & 74   & 99  & 144 & -   \\
      \cline{2-12}
          & 3   & 21   & 22  & 40   & 50  & 81   & 102  & 155  & 279 & -   & -   \\
      \cline{2-12}
          & 4   & 27   & 34  & 59   & 99  & 141  & 237  & 1114 & -   & -   & -   \\
      \cline{2-12}
          & 5   & 22   & 94  & 125  & 335 & \TO   & 1406 & -    & -   & -   & -   \\
      \cline{2-12}
          & 6   & 36   & 132 & 2966 & 844 & 2457 & -    & -    & -   & -   & -   \\
      \cline{2-12}
          & 7   & 83   & 487 & \TO   & \TO  & -    & -    & -    & -   & -   & -   \\
      \cline{2-12}
          & 8   & 298  & \TO  & \TO   & -   & -    & -    & -    & -   & -   & -   \\
      \cline{2-12}
          & 9   & 1428 & \TO  & -    & -   & -    & -    & -    & -   & -   & -   \\
      \cline{2-12}
          & 10   & \TO   & -   & -    & -   & -    & -    & -    & -   & -   & -   \\
      \cline{2-12}
  \end{tabular}
  \caption{Benchmark of full proof computation time for $\OMH(1)$ implementation. Times are in seconds with a timeout (\TO) limit of one hour. Dashes ('-') denote no benchmark was run.}
  \label{fig:benchmark}
\end{figure}

@@

\begin{figure}
  \centering
  \begin{tabular}{|p{2.25cm}|p{2.25cm}|p{2.25cm}|p{2.25cm}|p{2.25cm}|}
  \hline
                                  & Transition systems (7 total)  &Definitions\linebreak (58 total) & Invariants\linebreak (11 total) & Invariant classes (5 total) \\
  \hline \hline
  No Fault Model                  & receivers & 8 modified & 2 modified & ASM, faults \\
  \hline
  Omissive Asymmetric Faults       & none      & 4 modified & 2 modified & faults             \\
  \hline
  Time-Triggered Messaging        & source, relays, receivers & 3 new & 2 new, 3 modified & calendar, faults \\
  \hline
  Mid-Value Selection & receivers & 4 new, 3 modified & 2 modified & ASM, voting \\
  \hline
  \end{tabular}
  \caption{Refactoring effort for protocol modifications, measured by which portions of the model have to be modified.}
  \label{fig:effort}
\end{figure}

>>>>

\newboolean{submission}  %set to true for the submission version
%\setboolean{submission}{false}
\setboolean{submission}{true}
\ifthenelse
{\boolean{submission}}
{ %
  \newcommand{\lee}[1]{ } %
  \newcommand{\ben}[1]{ } %
  % Put your own TODO macros here and below
} %hide todo
{ %
  \newcommand{\lee}[1]{ {\color{blue}$<$lee: #1$>$} } %
  \newcommand{\ben}[1]{ {\color{purple}$<$ben: #1$>$} } %
  \usepackage{hyperref} %
  \usepackage[inline]{showlabels} %
}

@@

\author{Benjamin F Jones\inst{1} \and Lee Pike\inst{2}}

@@

\lee{Ben will redo. I'd just remove the Gurevich reference}
\ben{I still need to do this!}

@@

% \begin{example}
% Consider a simple distributed system involving 3 processors: $S$, $L_1$, and $L_2$. The purpose of the system is for $S$ to broadcast a message to $L_1$ and $L_2$ after which they perform a computation on the message and update their local state with the result. One might model this system as a transition system whose state is the product of the local processor states along with the state of the environment and any globally important phenomena such as time.
% 
% \begin{figure}
% \centering
% \begin{tikzpicture}[->,>=stealth',shorten >=1pt,auto,node distance=3cm,
%   thick,main node/.style={circle,fill=blue!20,draw,
%   font=\sffamily\Large\bfseries,minimum size=8mm}]
% 
%   \node[main node] (S) at (0,0) {$S$};
%   \node[main node] (L1) at (2,1) {$L_1$};
%   \node[main node] (L2) at (2,-1) {$L_2$};
% 
%   \path[every node/.style={font=\sffamily\small,
%           fill=white,inner sep=1pt}]
%     (S)  edge (L1)
%          edge (L2);
% 
% \end{tikzpicture}
% \label{fig:exa-trans}
% \caption{Simple message passing system}
% \end{figure}

Intuitively, there are subsets of the overall state space which are important to understanding the evolution of such a system. For example, let $A_1$ denote the subset of states in which the processor $S$ has not sent a message, let $A_2$ be the subset in which $S$ has broadcast a message, and let $A_3$ be the subset in which $L_1$ and $L_2$ have each received and processed the message. The union $A_1 \cup A_2 \cup A_3$ covers the set of reachable states of the system and defines an ASM that is useful for reasoning about its dynamics and understanding counter-example traces.

% \end{example}

\section{Putting it Together: Compositional Modeling and Verification for Oral Messages}\label{sec:byz}

@@

\begin{figure}
  \centering
  \begin{tabular}{c|c||c|c|c|c|c|c|c|c|c|c|}
      \multicolumn{1}{r}{} & \multicolumn{1}{r}{} & \multicolumn{10}{c}{Receivers} \\
      \cline{2-12}
          &     & 1  &  2  &   3  &  4  &   5  &  6   &  7   &  8  &  9  &  10 \\
      \cline{2-12} \cline{2-12}
          \multirow{10}{*}{Relays\phantom{x}} &
            1   & 7    & 9   & 12   & 15  & 21   & 25   & 32   & 40  & 54  & 74  \\
      \cline{2-12}
          & 2   & 17   & 14  & 21   & 30  & 42   & 53   & 74   & 99  & 144 & -   \\
      \cline{2-12}
          & 3   & 21   & 22  & 40   & 50  & 81   & 102  & 155  & 279 & -   & -   \\
      \cline{2-12}
          & 4   & 27   & 34  & 59   & 99  & 141  & 237  & 1114 & -   & -   & -   \\
      \cline{2-12}
          & 5   & 22   & 94  & 125  & 335 & \TO   & 1406 & -    & -   & -   & -   \\
      \cline{2-12}
          & 6   & 36   & 132 & 2966 & 844 & 2457 & -    & -    & -   & -   & -   \\
      \cline{2-12}
          & 7   & 83   & 487 & \TO   & \TO  & -    & -    & -    & -   & -   & -   \\
      \cline{2-12}
          & 8   & 298  & \TO  & \TO   & -   & -    & -    & -    & -   & -   & -   \\
      \cline{2-12}
          & 9   & 1428 & \TO  & -    & -   & -    & -    & -    & -   & -   & -   \\
      \cline{2-12}
          & 10   & \TO   & -   & -    & -   & -    & -    & -    & -   & -   & -   \\
      \cline{2-12}
  \end{tabular}
  \caption{Benchmark of full proof computation time for $\OMH(1)$ implementation. Times are in seconds with a timeout (\TO) limit of one hour. Dashes ('-') denote no benchmark was run.}
  \label{fig:benchmark}
\end{figure}

@@

\begin{figure}
  \centering
  \begin{tabular}{|p{2.25cm}|p{2.25cm}|p{2.25cm}|p{2.25cm}|p{2.25cm}|}
  \hline
                                  & Transition systems (7 total)  &Definitions\linebreak (58 total) & Invariants\linebreak (11 total) & Invariant classes (5 total) \\
  \hline \hline
  No Fault Model                  & receivers & 8 modified & 2 modified & ASM, faults \\
  \hline
  Omissive Asymmetric Faults       & none      & 4 modified & 2 modified & faults             \\
  \hline
  Time-Triggered Messaging        & source, relays, receivers, ASM & 3 new & 2 modified, 3 modified & calendar, faults \\
  \hline
  Mid-Value Selection & receivers & 4 new, 3 modified & 2 modified & ASM, voting \\
  \hline
  \end{tabular}
  \caption{Refactoring effort for protocol modifications, measured by which portions of the model have to be modified.}
  \label{fig:effort}
\end{figure}
